---
title: "BARI_split_plot_desig"
author: "MD. NEAZ ALI"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import Libraries
```{r}
library(lme4)
library(lmerTest)
library(emmeans)
library(ggplot2)
library(agricolae)
library(dplyr)
```

# Import Data
```{r}
# Data Import
df <- read.csv("./split_plot.csv", header = T)
head(df)
```
```{r}
cat("Dimension of the dataframe",dim(df))
cat("Str: ", str(df))
```

## Data Cleanings

```{r}
df$Rep <- as.factor(df$Rep)
df$Tillage <- as.factor(df$Tillage)
df$Organic_Manure <- as.factor(df$Organic_Manure)
table(df$Rep)
```

```{r}
table(df$Tillage)
```
```{r}
table(df$Organic_Manure)
```
# Group mean and sd

```{r}
df_summary <- df %>%
  group_by(Tillage) %>%
  summarise(
    Mean = mean(Yield, na.rm = TRUE),
    SD = sd(Yield, na.rm = TRUE),
    MAX = max(Yield, na.rm = TRUE),
    Min = min(Yield, na.rm = TRUE)
  )

# Print the summary
print(df_summary)
```
```{r}
# Create the aesthetic bar plot with error bars
ggplot(df_summary, aes(x = Tillage, y = Mean, fill = Tillage)) +
  geom_bar(stat = "identity", position = "dodge", show.legend = FALSE) +  # Create bars
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2) +  # Add error bars
  labs(
    title = "Group Means with Standard Deviation",
    x = "Group",
    y = "Mean Yield",
    caption = "Error bars represent standard deviation"
  ) +
  theme_minimal() +  # Use minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for clarity
    text = element_text(size = 14)  # Increase text size for better readability
  )
```
```{r}
df_summary_rep <- df %>%
  group_by(Rep) %>%
  summarise(
    Mean = mean(Yield, na.rm = TRUE),
    SD = sd(Yield, na.rm = TRUE),
    MAX = max(Yield, na.rm = TRUE),
    MIN = min(Yield, na.rm = TRUE)
  )

# Print the summary
print(df_summary_rep)
```
```{r}
# Create the aesthetic bar plot with error bars
ggplot(df_summary_rep, aes(x = Rep, y = Mean, fill = Rep)) +
  geom_bar(stat = "identity", position = "dodge", show.legend = FALSE) +  # Create bars
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2) +  # Add error bars
  labs(
    title = "Group Means with Standard Deviation",
    x = "Group",
    y = "Mean Yield",
    caption = "Error bars represent standard deviation"
  ) +
  theme_minimal() +  # Use minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for clarity
    text = element_text(size = 14)  # Increase text size for better readability
  )
```

```{r}
df_summary_manure <- df %>%
  group_by(Organic_Manure) %>%
  summarise(
    Mean = mean(Yield, na.rm = TRUE),
    SD = sd(Yield, na.rm = TRUE),
    MAX = max(Yield, na.rm = TRUE),
    MIN = min(Yield, na.rm = TRUE)
  )

# Print the summary
print(df_summary_manure)
```

```{r}
# Create the aesthetic bar plot with error bars
ggplot(df_summary_manure, aes(x = Organic_Manure, y = Mean, fill = Organic_Manure)) +
  geom_bar(stat = "identity", position = "dodge", show.legend = FALSE) +  # Create bars
  geom_errorbar(aes(ymin = Mean - SD, ymax = Mean + SD), width = 0.2) +  # Add error bars
  labs(
    title = "Group Means with Standard Deviation",
    x = "Group",
    y = "Mean Yield",
    caption = "Error bars represent standard deviation"
  ) +
  theme_minimal() +  # Use minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for clarity
    text = element_text(size = 14)  # Increase text size for better readability
  )
```

# ANOVA
```{r}
model <- aov(Yield ~ Tillage * Organic_Manure + Error(Rep/Tillage), data = df)
summary(model)
```

# Post Hoc Tests

### DMRMT Test
```{r}
model2 <- aov(Yield~Tillage+Organic_Manure+Rep, data = df)
duncan_result <- duncan.test(model2, "Organic_Manure", group = TRUE)

# Print the results
print(duncan_result)
```
### LSD Test
```{r}
# model2 <- aov(Yield~Tillage+Organic_Manure+Rep, data = df)
lsd_result <- LSD.test(model2, "Organic_Manure", group = TRUE)

# Print the results
print(lsd_result)
```

### HSD Test
```{r}
hsd_result <- HSD.test(model2, "Organic_Manure", group = TRUE)

# Print the results
print(hsd_result)
```




# Model Diagnostics

### Boxplot of residuals
```{r}
residuals_data <- residuals(model2)

# Create a data frame for ggplot
residuals_df <- data.frame(residuals = residuals_data)

# Plotting the boxplot using ggplot
ggplot(residuals_df, aes(x = 1, y = residuals)) +
  geom_boxplot(fill = "green", color = "black", outlier.colour = "red", outlier.size = 3) +
  labs(title = "Boxplot of Residuals", x = "Residuals", y = "Value") +
  theme_minimal() +
  theme(axis.text.x = element_blank(),  # Hide x-axis label (since we have only one category)
        axis.ticks.x = element_blank())  # Hide x-axis ticks
```


### Shapiro -Wilks Test
```{r}
residuals <- residuals(model2)

# Q-Q plot
qqnorm(residuals)
qqline(residuals, col = "red")

# Shapiro-Wilk test for normality
shapiro.test(residuals)
```


# Fit the Mixed-Effect Model

```{r}
model <- lmer(Yield ~ Tillage + Organic_Manure + Tillage * Organic_Manure + (1 | Rep), data = df)# Rep is random
summary(model)
```
```{r}
anova(model)
```

# Post_Hoc Tests
```{r}
emm_tillage <- emmeans(model, ~ Tillage)
emm_manure <- emmeans(model, ~ Organic_Manure)
emm_interaction <- emmeans(model, ~ Tillage * Organic_Manure)
```

```{r}
print(summary(emm_tillage))
print(summary(emm_manure))
print(summary(emm_interaction))
```

# Pairwise
```{r}
pairs(emm_tillage)
```
```{r}
pairs(emm_manure)
```

```{r}
pairs(emm_interaction)
```


# Visualize Results

```{r}
ggplot(df, aes(x = Tillage, y = Yield, fill = Organic_Manure)) +
  geom_boxplot() +
  labs(title = "Interaction between Tillage and Organic Manure on Yield") +
  theme_minimal()
```

# Diagnostic

## Residual Plot

### Residuals vs Fitted Plot
```{r}
residuals_model  <- residuals(model)
summary(residuals_model)
```

```{r}
# Extract residuals and fitted values
fitted_values <- fitted(model)
residuals_values <- residuals(model)

# Create the ggplot
ggplot(data.frame(Fitted = fitted_values, Residuals = residuals_values), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue", alpha = 0.7) + 
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
```

## Influential Points
```{r}
# Check for Cook's distance
cooks_d <- cooks.distance(model)
plot(cooks_d, main = "Cook's Distance", ylab = "Cook's Distance")
abline(h = 4 / length(cooks_d), col = "red")  # Threshold for influential points
```
```{r}
library(car)
vif(model)
```

## Q-Q Plot

```{r}
ggplot(data.frame(Residuals = residuals_values), aes(sample = Residuals)) +
  stat_qq() +
  stat_qq_line(color = "red") +
  labs(title = "Normal Q-Q Plot") +
  theme_minimal()
```

## Scale - Location Plot

None


## Cooks Distance Plot
```{r}
# Calculate Cook's Distance
cooks_distance <- cooks.distance(model)

# Create a Cook's distance plot
ggplot(data.frame(Index = seq_along(cooks_distance), CooksDistance = cooks_distance), aes(x = Index, y = CooksDistance)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 4 / length(cooks_distance), color = "red", linetype = "dashed") +
  labs(title = "Cook's Distance", x = "Index", y = "Cook's Distance") +
  theme_minimal()
```

## Autocorrelation of Residuals (ACF Plot)
```{r}
# Plot the autocorrelation of residuals
acf(residuals(model), main = "Autocorrelation of Residuals", col = "blue", lwd = 2)
```
## Fitted Values vs Observed Values

```{r}
# Create a fitted vs observed plot
ggplot(data.frame(Observed = df$Yield, Fitted = fitted_values), aes(x = Observed, y = Fitted)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Fitted vs Observed", x = "Observed Values", y = "Fitted Values") +
  theme_minimal()

```

```{r}
# rstandard(model) # Not Running
```


```{r}
library(gridExtra)

fitted_values <- fitted(model)
residual_sd <- sd(residuals_model)
std_residuals <- residuals_model/residual_sd

# Create individual plots
p1 <- ggplot(data.frame(Fitted = fitted_values, Residuals = residuals_values), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "blue", alpha = 0.7) + 
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()

p2 <- ggplot(data.frame(Fitted = fitted_values, StdResiduals = std_residuals), aes(x = Fitted, y = StdResiduals)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Scale-Location (Standardized Residuals vs Fitted)", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

p3 <- ggplot(data.frame(Fitted = fitted_values, StdResiduals = std_residuals), aes(x = Fitted, y = StdResiduals)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "red", linetype = "dashed") +
  labs(title = "Scale-Location", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

p4 <- ggplot(data.frame(Index = seq_along(cooks_distance), CooksDistance = cooks_distance), aes(x = Index, y = CooksDistance)) +
  geom_bar(stat = "identity", fill = "steelblue", alpha = 0.7) +
  geom_hline(yintercept = 4 / length(cooks_distance), color = "red", linetype = "dashed") +
  labs(title = "Cook's Distance", x = "Index", y = "Cook's Distance") +
  theme_minimal()

# Combine all plots into a grid
grid.arrange(p1, p2, p3, p4, ncol = 2)
```


##  Basic Boxplot (Yield by Tillage and Manure)

None

## 2. Boxplot with Facets (by Replications)

```{r}
# Check the unique levels of Organic_Manure
unique_manure_levels <- levels(df$Organic_Manure)

# Define the number of colors needed based on the levels
n_levels <- length(unique_manure_levels)
color_palette <- c("lightblue", "lightgreen", "lightpink", "lightyellow")  # Adjust colors if needed

# Generate the plot with adjusted color palette
ggplot(df, aes(x = Tillage, y = Yield, fill = Organic_Manure)) +
  geom_boxplot() +
  facet_wrap(~ Rep, scales = "free_y") +  # Separate boxplots for each Replication
  labs(title = "Boxplot of Yield by Tillage, Manure, and Replication", x = "Tillage", y = "Yield") +
  scale_fill_manual(values = color_palette[1:n_levels]) +  # Use the correct number of colors
  theme_minimal() +
  theme(legend.position = "top")
```
```{r}
ggplot(df, aes(x = Tillage, y = Yield, fill = Tillage)) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Set3") +  # Use a color palette for the box fill
  labs(title = "Boxplot of Yield by Tillage", 
       x = "Tillage", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```
```{r}
# Aesthetic Boxplot of Yield by Organic_Manure
ggplot(df, aes(x = Organic_Manure, y = Yield, fill = Organic_Manure)) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Pastel1") +  # Use a light, pastel color palette
  labs(title = "Boxplot of Yield by Organic Manure", 
       x = "Organic Manure", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```
```{r}
# Aesthetic Boxplot of Yield by Tillage and Organic Manure
ggplot(df, aes(x = interaction(Tillage, Organic_Manure), y = Yield, fill = interaction(Tillage, Organic_Manure))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Set3") +  # Use a color palette for the interaction levels
  labs(title = "Boxplot of Yield by Tillage and Organic Manure", 
       x = "Tillage and Organic Manure", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```
```{r}
# Aesthetic Boxplot of Yield by Replication (Rep)
ggplot(df, aes(x = Rep, y = Yield, fill = Rep)) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Set1") +  # Use a color palette for Replication groups
  labs(title = "Boxplot of Yield by Replication", 
       x = "Replication", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```
```{r}
# Aesthetic Boxplot of Yield by Rep and Tillage
ggplot(df, aes(x = interaction(Rep, Tillage), y = Yield, fill = interaction(Rep, Tillage))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Set3") +  # Use a color palette for the interaction levels
  labs(title = "Boxplot of Yield by Replication and Tillage", 
       x = "Replication and Tillage", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```
```{r}
# Aesthetic Boxplot of Yield by Rep and Organic Manure
ggplot(df, aes(x = interaction(Rep, Organic_Manure), y = Yield, fill = interaction(Rep, Organic_Manure))) +
  geom_boxplot(outlier.shape = 16, outlier.size = 2, outlier.colour = "red") +  # Customize outliers
  scale_fill_brewer(palette = "Set3") +  # Use a color palette for the interaction levels
  labs(title = "Boxplot of Yield by Replication and Organic Manure", 
       x = "Replication and Organic Manure", 
       y = "Yield") +
  theme_minimal() +  # Clean minimal theme
  theme(
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),  # Center and bold title
    axis.title = element_text(size = 14),  # Axis title size
    axis.text = element_text(size = 12),  # Axis text size
    legend.position = "none"  # Remove legend
  )
```

# Histograms

## Basic Density Curve
```{r}
ggplot(df, aes(x = Yield)) +
  geom_histogram(aes(y = ..density..), 
                 bins = 30, 
                 fill = "green", 
                 color = "black", 
                 alpha = 0.7) +  # Adjust histogram transparency
  geom_density(color = "black", size = 1.2) +  # Overlay density curve
  labs(title = "Histogram with Density Curve for Yield", 
       x = "Yield", 
       y = "Density") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))  # Title customization
```
## Histogram with Density Curve for Yield by Tillage:
```{r}
ggplot(df, aes(x = Yield, fill = Tillage)) +
  geom_histogram(aes(y = ..density..), 
                 bins = 30, 
                 color = "black", 
                 alpha = 0.6) +
  geom_density(aes(color = Tillage), size = 1.2) +  # Separate density curves by Tillage
  labs(title = "Histogram with Density Curve by Tillage", 
       x = "Yield", 
       y = "Density") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))

```
## Histogram with Density Curve for Yield by Organic Manure:
```{r}
ggplot(df, aes(x = Yield, fill = Organic_Manure)) +
  geom_histogram(aes(y = ..density..), 
                 bins = 30, 
                 color = "black", 
                 alpha = 0.6) +
  geom_density(aes(color = Organic_Manure), size = 1.2) +  # Separate density curves by Organic Manure
  labs(title = "Histogram with Density Curve by Organic Manure", 
       x = "Yield", 
       y = "Density") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```
## Histogram with Density Curve for Yield by Replication:
```{r}
ggplot(df, aes(x = Yield, fill = Rep)) +
  geom_histogram(aes(y = ..density..), 
                 bins = 30, 
                 color = "black", 
                 alpha = 0.6) +
  geom_density(aes(color = Rep), size = 1.2) +  # Separate density curves by Replication
  labs(title = "Histogram with Density Curve by Replication", 
       x = "Yield", 
       y = "Density") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```
## Facet Grid Histogram with Density Curve for Yield by Tillage and Organic Manure:
```{r}
ggplot(df, aes(x = Yield, fill = Tillage)) +
  geom_histogram(aes(y = ..density..), 
                 bins = 30, 
                 color = "black", 
                 alpha = 0.6) +
  geom_density(aes(color = Tillage), size = 1.2) +  # Separate density curves by Tillage
  facet_wrap(~ Organic_Manure) +  # Facet by Organic Manure
  labs(title = "Histogram with Density Curve by Tillage and Organic Manure", 
       x = "Yield", 
       y = "Density") +
  theme_minimal() + 
  theme(plot.title = element_text(size = 16, face = "bold", hjust = 0.5))
```

